---
title: "R First Look"
output: 
  pdf_document: default
  html_notebook: default
---

##Loading the data
Our data will either be an Excel file or a comma-separated values (.csv) file. These are very common and readable by many different programs. For most of the R work in this class we will make use of ".csv" files because they make it easier to call the data directly from online. There are packages for loading data into R that provide commands. Much of what we will use this semester is located in the \em{tidyverse}. This is a collection of the most popular collections of code and commands in R. It includes collections for reading data, data analysis, and data visualization.

In addition, for now, to make sense of the date information we will make use of a package called \em{tsibble}. This is useful for time/date type visualizations.^[I gave an explanation of the origin of this term in class.] The use of the url points to the data set used for this introduction to R. It is downloaded directly from github into R saving us steps.^[This would also be a useful feature pointing, for example, to an internal company server that will allow someone to provide a file to others without having to provide, simultaneously, the data set.]

```{r}
library(tidyverse)
library(tsibble)
urlfile="https://raw.githubusercontent.com/flynnecon/datavizf24/refs/heads/main/online_monthly_R.csv"
```
#Getting the data read into R and then made into our time series analysis format. Note we do not need a training or test data division with the fact we are looking at a forecast 1 period ahead.
```{r  data work}
data <- read_csv(url(urlfile))
data <- data[-c(1)]
data$month <- seq(as.Date("2018-06-01"), as.Date("2024-07-01"), by="1 month")
data_tsbl <- data |> mutate(date=yearmonth(month)) |> as_tsibble(index=date)
```

```{r scatter}
data_tsbl |> ggplot(aes(x=date,y=Handle)) + geom_point(color="darkgreen")
```

```{r another scatter}
ggplot(data=data_tsbl,aes(x=date,y=Handle/1000000000)) + geom_line(color="blue")
```

```{r combination}
ggplot(data=data_tsbl, aes(x=date,y=Handle)) + geom_point(color="darkgreen") + geom_line(color="blue") + theme_bw()
```
